---
title: "Full rating and directory data inspection"
author: "Meng Ye"
date: "3/28/2025"
output:
  pdf_document: default
  html_document: default
  word_document: default
---

## Introduction and Setup 

This file contains the exploration of the two full CQC data files "cqc_full" and "raw_directory".

### Load packages 

```{r load package, message=FALSE, warning=FALSE}
library(tidyverse) # package for data cleaning and plotting
library(jsonlite) # to read json data
library(readODS)  # read ODS file 
library(readxl)   # read excel
library(janitor) # clean variable names
library(here) # manage directory
library(gt)
library(writexl)
set.seed(5432)
```


### Import Raw Data 


```{r import data, message=FALSE, warning=FALSE}
# Raw CQC full data (with unfriendly column names) to over sampling CICs
rating_raw <- read_csv(here("data", "raw_data", "cqc_full.csv"))
 
# import charities commission data
charities <- read_csv(here("data", "raw_data", "charities.csv"))

# manually coded (cross-checking against Companies House registration record) individuals and partnerships providers data
directory_raw <- read_excel(here("data", "raw_data", "full_directory.xlsx"))
```


## Data insepection

```{r}
directory_rename <- directory_raw |>
  clean_names()
```


```{r}
# Count records per region and local authority
distribution_table <- directory_rename %>%
  group_by(location_region, location_local_authority) %>%
  summarise(n = n(), .groups = "drop")  # Count observations

# Print the table (sorted by region and count)
distribution_table %>%
  arrange(location_region, desc(n)) %>% gt()
```




```{r}
n_distinct(rating_raw$`Provider ID`)
```

```{r}
n_distinct(directory_raw$`Provider ID`)
```
```{r}
na_summary_tidy <- directory_rename %>%
  select(starts_with("service_user_band_")) %>%   # Select target columns
  summarise(across(everything(), ~sum(is.na(.x)))) %>%  # Count NAs
  pivot_longer(
    everything(),
    names_to = "column",
    values_to = "na_count"
  ) %>%
  mutate(
    na_ratio = na_count / nrow(directory_rename),  # Calculate ratio
    na_percent = round(na_ratio * 100, 1)          # Optional: add %
  ) %>%
  arrange(desc(na_count))                          # Sort by highest NA count

# Print the tidy summary
na_summary_tidy
```

```{r message=FALSE, warning=FALSE}
# Save as Excel
write_xlsx(na_summary_tidy, here("data", "derived_data", "na_summary_tidy.xlsx"))

```


```{r}
library(dplyr)

# Count rows where ALL columns are NA
all_na_count <- directory_rename %>%
  filter(if_all(everything(), is.na)) %>%  # Keep rows where all columns are NA
  summarise(all_na_count = n())            # Count those rows

# Print the result
all_na_count
```


No missing values for the question

### Clean and filter out Gov and CIC providers

```{r rename and filter, message=FALSE, warning=FALSE}
### Rename the columns 

locations_renamed <- rating_raw %>% 
  rename(primary_cat = `Location Primary Inspection Category`,
         region = `Location Region`,
         domain = `Domain`, # named to "Domain" later
         rating = `Latest Rating`,
 #        inherited = `Inherited Rating (Y/N)`, # no such a column in 2017 data
         provider_name = `Provider Name`,
         publication_date = `Publication Date`,
         address = `Location Street Address`) 
```



### Standardize the provider name

```{r sample std}
locations_std <- locations_renamed %>% 
# remove dots
  mutate(std_name = str_replace_all(provider_name, "\\.", "")) %>% 
# capitalize all
  mutate(std_name = str_to_upper(std_name)) %>%
# standardize to LTD
  mutate(std_name = str_replace_all(std_name, "LIMITED", "LTD")) %>% 
# standardize to CIC
  mutate(std_name = str_replace_all(std_name, "COMMUNITY INTEREST COMPANY", "CIC"))
```


### filtering out CIC and GOV providers

```{r}
## building the string detect conditions
keep_terms <- c("CIC", "NHS ", "FOUNDATION TRUST", "COUNCIL", " CC ", " MDC ", " DC ", " BC ", "BOROUGH")
keep_filter <- paste0("\\b(", paste(keep_terms, collapse = "|"), ")\\b")
```

###

```{r}
## filtering out CICs and GOVs
locations_coded <- locations_std %>% 
  filter(str_detect(std_name, keep_filter)) %>% 
#  filter(!str_detect(std_name, "BOROUGH GREEN\\s|BOROUGH ROAD\\s")) %>% (save for false positives)
  ## add the form and level columns
  mutate(form = ifelse(str_detect(std_name, "CIC"), "CIC", "GOV")) 
nrow(locations_coded)
```
```{r}
#filter out the government list

gov_list <- locations_coded %>% 
  filter(form == "GOV") %>% 
  select(provider_name) %>% 
  group_by(provider_name) %>% 
  slice(1)
```


```{r message=FALSE, warning=FALSE}

# Save as Excel
write_xlsx(gov_list, here("data", "derived_data", "gov_list.xlsx"))

```


----------separate line ------------

Derive CQC provider list

```{r}
cqc_provider <- directory_rename %>% 
  select(provider_name, provider_companies_house_number, provider_charity_number, provider_ownership_type) %>% 
  group_by(provider_name) %>% 
  slice(1)

```

```{r}
# Save as Excel
write_xlsx(cqc_provider, here("data", "derived_data", "cqc_provider.xlsx"))
```


----------separate line ------------

### 1a. Rename coded sample data frame columns to be more R-friendly 

```{r rename sample}
socialcare_renamed <- socialcare_raw %>% 
  rename(primary_cat = `Location Primary Inspection Category`,
         region = `Location Region`,
         domain = Domain,
         rating = latest_rating,
         inherited = `Inherited Rating (Y/N)`,
         provider_name = `Provider Name`,
         form_num = legal_form,
         publication_date = `Publication Date`) %>% 
  select(- `...1`, -`Location Type`) %>% 
  filter(location_type == "Social Care Org")
```

```{r}
head(socialcare_renamed)
```



### 1b. Standardize coded sample data frame's org name expression

```{r sample std}
socialcare_renamed <- socialcare_renamed %>% 
# remove dots
  mutate(std_name = str_replace_all(provider_name, "\\.", "")) %>% 
# capitalize all
  mutate(std_name = str_to_upper(std_name)) %>%
# standardize to LTD
  mutate(std_name = str_replace_all(std_name, "LIMITED", "LTD")) %>% 
# standardize to CIC
  mutate(std_name = str_replace_all(std_name, "COMMUNITY INTEREST COMPANY", "CIC"))
```


### 2a. Rename full cqc data frame columns to be more R-friendly 

```{r rename cqc full data set}
cqc_renamed <- cqc_raw %>% 
  rename(primary_cat = `Location Primary Inspection Category`,
         region = `Location Region`,
         domain = Domain,
         rating = `Latest Rating`,
         inherited = `Inherited Rating (Y/N)`,
         provider_name = `Provider Name`,
         location_id = `Location ID`,
         location_type = `Location Type...26`,
         publication_date = `Publication Date`) %>% 
  select(-`Location Type...5`, -`Legal Form`) %>% 
  filter(location_type == "Social Care Org")


```


### 2b. Standardize coded sample data frame's org name expression

```{r cqc std}
cqc_renamed <- cqc_renamed %>% 
# remove dots
  mutate(std_name = str_replace_all(provider_name, "\\.", "")) %>% 
# capitalize all
  mutate(std_name = str_to_upper(std_name)) %>%
# standardize to LTD
  mutate(std_name = str_replace_all(std_name, "LIMITED", "LTD")) %>% 
# standardize to CIC
  mutate(std_name = str_replace_all(std_name, "COMMUNITY INTEREST COMPANY", "CIC")) %>% 
# standardize to CIO
  mutate(std_name = str_replace_all(std_name, "CHARITABLE INCORPORATED ORGANISATION", "CIO")) 
```


### 3. Standardize CIC list org name expression 

```{r cic full list std}
cic_renamed <- cic_recoded %>% 
  select(-`...1`, - legal_name, cic_number)  %>% 
# remove dots
  mutate(std_name = str_replace_all(provider_name, "\\.", "")) %>% 
# capitalize all
  mutate(std_name = str_to_upper(std_name)) %>%
# standardize to LTD
  mutate(std_name = str_replace_all(std_name, "LIMITED", "LTD")) %>% 
# standardize to CIC
  mutate(std_name = str_replace_all(std_name, "COMMUNITY INTEREST COMPANY", "CIC")) %>% 
# standardize to CIO
  mutate(std_name = str_replace_all(std_name, "CHARITABLE INCORPORATED ORGANISATION", "CIO")) %>% 
# delete unnecessary columns to make sure combining row align better later
  select(- cic_number, - provider_name)
```


### 4a. Derive a smaller sub set (select necessary columns) of Charity Commision data

```{r CC subset}
charity_selected <- charities %>% 
# filter out only active charities
  filter(charity_registration_status == "Registered") %>% 
# select only potentially useful columns 
  select(charity_number = registered_charity_number,
         charity_name, charity_type,
         charity_registration_date = date_of_registration,
         charity_company_number = charity_company_registration_number)
```


### 4b. Standardize Charity Commision org name expression 

```{r CC std}
charity_renamed <- charity_selected %>% 
# remove dots
  mutate(std_name = str_replace_all(charity_name, "\\.", "")) %>% 
# capitalize all
  mutate(std_name = str_to_upper(std_name)) %>%
# standardize to LTD
  mutate(std_name = str_replace_all(std_name, "LIMITED", "LTD")) %>% 
# standardize to CIC
  mutate(std_name = str_replace_all(std_name, "COMMUNITY INTEREST COMPANY", "CIC")) %>% 
# standardize to CIO
  mutate(std_name = str_replace_all(std_name, "CHARITABLE INCORPORATED ORGANISATION", "CIO")) 
```


### 5.Standardize individual/partnership provider name expression 

Based on Dr. Kerlin's code, this is what I did:


1.	Leave it as the same type individual/partnerships for blank rows
2.  Leave it as the same type individual/partnerships for rows with No-No for both Charity Commission and Company House
3.	Recode those have limited by shares Company House records as “for-profit”. 
4.	Recode “LLP” to for-profit 
5.  Recode company limited by guarantee to "NA", because we are not sure if it is nonprofit or not
(See https://en.wikipedia.org/wiki/Private_company_limited_by_guarantee, especially under Commercial enterprises)
4.	Recode to “NA” for those marked with several possibilities 


```{r rename indivial and partnership}
small_renamed <- small %>% 
  select(provider_name, small_form_num) %>% 
  
# remove dots
  mutate(std_name = str_replace_all(provider_name, "\\.", "")) %>% 
# capitalize all
  mutate(std_name = str_to_upper(std_name)) %>%
# standardize to LTD
  mutate(std_name = str_replace_all(std_name, "LIMITED", "LTD")) %>% 
# standardize to CIC
  mutate(std_name = str_replace_all(std_name, "COMMUNITY INTEREST COMPANY", "CIC")) %>% 
# standardize to CIO
  mutate(std_name = str_replace_all(std_name, "CHARITABLE INCORPORATED ORGANISATION", "CIO")) %>% 
  select(-provider_name)
```



## Data cleanning - merging and recoding 

### 1. Use full CIC list and complete CQC data to derive full CIC data frame

```{r cic over sample data}
cic_over_sample <- cic_renamed %>% 
  left_join(cqc_renamed, by = "std_name") %>% 
# mark that the form is cic
  mutate(form_num = 4)
```

Now we check the merged data, there are 540 rows (for 57 organizations because of multiple locations and multiple rating domains) and 33 variables. 
```{r cic over sample data check}
head(cic_over_sample, 10)
```


### 2. Merge orginical sample with the full CIC sample 

First we carve out the CICs from the old samples 
```{r merge old sample and cic}
sample_no_cic <- socialcare_renamed %>% 
## leave out old cic orgs from the coded sample
  filter(form_num != 4)
```

Then we combine the sample without (old) CICs and the full CICs data

```{r add full cic}
sample_v2 <- bind_rows(sample_no_cic, cic_over_sample)
head(sample_v2)
```


### 3. Merge the updated sample v2 with indivial/partnership data

```{r merge in small}
# select fewer columns to merge
sample_v3 <- sample_v2 %>% 
# add data with refined coding
  left_join(small_renamed, by = "std_name") %>% 
# supersede old coding with new coding
  mutate(form_num = ifelse(form_num == 5, small_form_num, form_num)) %>% 
  select(- small_form_num)

tail(sample_v3)
```

### 4. Merge the updated sample v3 with Charity data

```{r}
# select charity std names and code the form type as "2".
charity_list <- charity_renamed %>% 
  select(std_name) %>% 
  mutate(charity_form_num = 2) %>% 
# Drop repetition
  group_by(std_name) %>% 
  slice(1) %>% 
  ungroup()

sample_v4 <- sample_v3 %>% 
# add data with refined coding
  left_join(charity_list, by = "std_name") %>% 
# supersede old coding with new coding
  mutate(form_num = ifelse(is.na(charity_form_num), form_num, charity_form_num)) %>% 
  select(- charity_form_num)

sample_new <- sample_v4%>% 
  mutate(index = 1:nrow(sample_v4))
```


## Check how much change was made


```{r distribution old name}
sample_new %>% 
  count(form_num)
```


Note:
                case_when(form_num == 1 ~ "FPO",
                          form_num == 2 ~ "NPO",
                          form_num == 3 ~ "GOV",
                          form_num == 4 ~ "CIC",
                          form_num == 5 ~ "IND",


## Save data

```{r eval = FALSE}
write_rds(sample_new, file = here("data", "derived_data", "sample_new.rds"))
write_csv(sample_new, file = here("data", "derived_data", "sample_new.csv"))
```





```{r}
count(sample_new, factor(form_num), domain)
```

```{r count-by-level}
sample_new %>% 
  mutate(overall = ifelse(domain == "Overall", 1, 0)) %>% 
  summarize(count_provider = n_distinct(provider_name),
            count_location = n_distinct(location_id),
            count_overall_rating = sum(overall),
            count_rating = n())
```

```{r count-by-form-level}
# counts of multi-level observations by the form types
# statistics corresponding to Table 1
sample_new %>% 
  mutate(overall = ifelse(domain == "Overall", 1, 0)) %>% 
  group_by(factor(form_num)) %>% 
  summarize(count_provider = n_distinct(provider_name),
            count_location = n_distinct(location_id),
            count_overall_rating = sum(overall),
            count_rating = n())
```


Note:
                case_when(form_num == 1 ~ "FPO",
                          form_num == 2 ~ "NPO",
                          form_num == 3 ~ "GOV",
                          form_num == 4 ~ "CIC",
                          form_num == 5 ~ "IND",
                          
                          
                          
                          
