---
title: "Full rating data cleaning"
author: "Meng Ye"
date: "4/22/2025"
output:
  pdf_document: default
  html_document: default
  word_document: default
---

## Introduction and Setup 

This file contains the exploration of the two full CQC data files "cqc_full" and "raw_directory".

### Load packages 

```{r load package, message=FALSE, warning=FALSE}
library(tidyverse) # package for data cleaning and plotting
library(jsonlite) # to read json data
library(readODS)  # read ODS file 
library(readxl)   # read excel
library(janitor) # clean variable names
library(here) # manage directory
library(gt)
library(writexl)
set.seed(5432)
```


### Import Raw Data 


```{r import data, message=FALSE, warning=FALSE}
# Raw CQC full data (with unfriendly column names) to over sampling CICs
rating_raw <- read_excel(here("data", "raw_data", "full_ratings.xlsx"))
 
# import charities commission data
charities <- read_csv(here("data", "raw_data", "charities.csv"))

# manually coded (cross-checking against Companies House registration record) individuals and partnerships providers data
directory_raw <- read_excel(here("data", "raw_data", "full_directory.xlsx"))
```


## Clean the location directory data-set

```{r}
directory_rename <- directory_raw |>
  clean_names()
```


```{r}
# select columns to use in the directory data set
directory_to_join <- directory_rename %>% 
  select(location_id, care_homes_beds, provider_companies_house_number, provider_charity_number,
         provider_ownership_type, starts_with("service_user_band")) %>% 
          mutate(across(
          # Target service user band columns
          starts_with("service_user_band"),  # <-- Verify this is the correct prefix!
          
          # Recode logic
          ~ case_when(
            .x == "Y" ~ TRUE,
            is.na(.x) ~ FALSE,
            TRUE      ~ FALSE  # Default for other values
          )))
```


## Clean the rating data-set

```{r}
# recode the column names}
rating_rename <- rating_raw |>
  clean_names() %>% 
  rename(inherited = inherited_rating_y_n,
         rating = latest_rating,
         primary_cat = location_primary_inspection_category) %>% 
  filter(location_type == "Social Care Org")
# print out the recoded col names
colnames(rating_rename)
```

## merge the two data sets


```{r rename and filter, message=FALSE, warning=FALSE}
rating_merged <- rating_rename %>% 
  left_join(directory_to_join, by = "location_id")
# print out the column names
colnames(rating_merged)
```

## coding the legal forms

### Standardize the provider name

```{r sample std}
rating_std <- rating_merged %>% 
# remove dots
  mutate(std_name = str_replace_all(provider_name, "\\.", "")) %>% 
# capitalize all
  mutate(std_name = str_to_upper(std_name)) %>%
# standardize to LTD
  mutate(std_name = str_replace_all(std_name, "LIMITED", "LTD")) %>% 
# standardize to CIC
  mutate(std_name = str_replace_all(std_name, "COMMUNITY INTEREST COMPANY", "CIC"))
```


### Code the GOV and CIC

Set the string detectors

```{r}
## building the string detect conditions
gov_terms <- c("NHS ", "FOUNDATION TRUST", "COUNCIL", " CC ", " MDC ", " DC ", " BC ", "BOROUGH")
gov_filter <- paste0("\\b(", paste(gov_terms, collapse = "|"), ")\\b")

cic_terms <- c("CIC")
cic_filter <- paste0("\\b(", paste(cic_terms, collapse = "|"), ")\\b")
```

###

```{r}
## filtering out CICs and GOVs
rating_detect <- rating_std %>% 
  mutate(
    form = case_when(
      str_detect(std_name, gov_filter) ~ "GOV",
      str_detect(std_name, cic_filter) ~ "CIC",
      TRUE ~ NA_character_)) %>% 
  mutate(
    form = case_when(
      # 1. Preserve existing CIC/GOV classifications
      form %in% c("CIC", "GOV") ~ form,
      is.na(form) & !is.na(provider_companies_house_number) ~ "FPO",
      TRUE ~ form)) %>% 
  mutate(
    form = ifelse(!is.na(provider_charity_number), "NPO", form)
  ) %>% 
  mutate(
    form = ifelse(provider_ownership_type %in% c("Partnership", "Individual"), "IND", form)
  )
    
```

```{r}
rating_detect %>%
  count(form, name = "n_rows")
```





###  Derive a smaller sub set (select necessary columns) of Charity Commision data

```{r CC subset}
charity_selected <- charities %>% 
# filter out only active charities
  filter(charity_registration_status == "Registered") %>% 
# select only potentially useful columns 
  select(charity_number = registered_charity_number,
         charity_name, charity_type,
         charity_registration_date = date_of_registration,
         charity_company_number = charity_company_registration_number)
```


### Standardize Charity Commision org name expression 

```{r CC std}
charity_renamed <- charity_selected %>% 
# remove dots
  mutate(std_name = str_replace_all(charity_name, "\\.", "")) %>% 
# capitalize all
  mutate(std_name = str_to_upper(std_name)) %>%
# standardize to LTD
  mutate(std_name = str_replace_all(std_name, "LIMITED", "LTD")) %>% 
# standardize to CIC
  mutate(std_name = str_replace_all(std_name, "COMMUNITY INTEREST COMPANY", "CIC")) %>% 
# standardize to CIO
  mutate(std_name = str_replace_all(std_name, "CHARITABLE INCORPORATED ORGANISATION", "CIO")) 
```



### update the charity list by merging in the charities commissions data

```{r}
# select charity std names and code the form type as "NPO".
charity_list <- charity_renamed %>% 
  select(std_name) %>% 
  mutate(charity_form = "NPO") %>% 
# Drop repetition
  group_by(std_name) %>% 
  slice(1) %>% 
  ungroup()

rating_clean <- rating_detect %>% 
# add data with refined coding
  left_join(charity_list, by = "std_name") %>% 
# supersede old coding with new coding
  mutate(form = ifelse(is.na(charity_form), form, charity_form)) %>% 
  select(- charity_form)
```


```{r}
rating_clean %>%
  count(form, name = "n_rows")
```



## Save data

```{r eval = FALSE}
write_rds(rating_clean, file = here("data", "derived_data", "rating_clean.rds"))
write_csv(rating_clean, file = here("data", "derived_data", "rating_clean.csv"))
```




