---
title: "Social Care CQC data analysis as submitted to PMR"
author: "Meng Ye"
date: "03/28/2024"
output:
  html_document: default
  pdf_document: default
  word_document: default
---

## Read Me

This is the data analysis RMarkdown file for the QCQ data after ISIRC 2022.

Mainly changes: incorporating the control of before and after (during) COVID

Please read the subtitles and notes added as normal text in this document. Blocks with darker backgrounds are code chunks, mostly with their outputs. I tried to add *#comments* within the chunks just before the code line to explain the purpose of the code line. Meanwhile I explain the purpose of the section in the texts, so that is where to find information to give you fuller pictures. 

```{r contr}
# environment setup to run ordered logit properly
options(contrasts = rep("contr.treatment", 2))
```


This chunk loads all the packages to use

```{r load package, message=FALSE, warning=FALSE}
#packages for ordered logit
library(ordinal) # package for ordinal logit regression
library(brant) # brant test for the parallel assumption for ordered logit
library(MASS) # models that work with the brant test

library(tidyverse) # package for data cleaning and plotting
library(readxl) # package for reading excel file
library(broom) # extracting model summary as data frame
library(modelsummary) # deriving model tables
library(scales) # label percent
library(lubridate) # working with dates
library(marginaleffects) # to calculate marginal effects
library(here) # manage directory
```



## Data Preparation

First, import the sampled and coded data set

```{r import data, message=FALSE, warning=FALSE}
#import the curated social care data file
#socialcare_raw <- read_csv("data/sample_new_cleaned.csv")
socialcare_new <-read_rds(here("data", "derived_data", "sample_new.rds"))
```

Assign orders to the ordinal level variables and name the organizational form in a reader-friendly way. 

```{r data cleaning}
#select relevant columns, rename and relabel 
socialcare <- socialcare_new %>% 
  # recode legal form types to be more readable / easier to present
  mutate(location_id = factor(location_id),
         form = case_when(form_num == 1 ~ "FPO",
                          form_num == 2 ~ "NPO",
                          form_num == 3 ~ "GOV",
                          form_num == 4 ~ "CIC",
                          form_num == 5 ~ "IND",
                          TRUE ~ NA_character_),
         inherited = ifelse(inherited == "Y", TRUE, FALSE),
         rating = recode(rating, 
                         "Insufficient evidence to rate" = "NA",
                         "Requires improvement" = "Req improv"),
         publication_date = dmy(publication_date)) %>% 
  # set the order of the values in the factors 
  mutate(form = fct_relevel(form, "FPO"),
         
         # assume the order of the ratings as follows but need to double check with the source 
         rating = ordered(rating, levels = c("Inadequate","Req improv", "Good", "Outstanding"))) %>% 
  
  # creating a new dummy variable for facility category
  mutate(category = case_when(
                    primary_cat == "Community based adult social care services" ~ "community",
                    primary_cat == "Residential social care" ~ "residential",
                    TRUE ~ as.character(NA)),
         # deriving year column and dummy variable for before_covid
         year = year(publication_date),
         during_covid = ifelse(year >= 2020, TRUE, FALSE),
         before_covid = ifelse(year <= 2019, TRUE, FALSE)) %>%

  # converting the ordinal variable to numerical 
  mutate(rating_num = case_when(rating == "Inadequate" ~ 1,
                                rating == "Req improv" ~ 2,
                                rating == "Good" ~ 3,
                                rating == "Outstanding" ~ 4)) %>% 
  # derive the rating dummy
  mutate(rating_higher = ifelse(rating_num > 2, 1, 0))

# show first several rows of the data set derived 
head(socialcare)

```


## Summerize at different Levels

```{r}
socialcare %>% 
  summarise(observation = n(),
            location = n_distinct(location_id),
            provider = n_distinct(provider_name))
```



```{r}
socialcare %>% 
  group_by(form) %>% 
  summarise(observation = n(),
            location = n_distinct(location_id),
            provider = n_distinct(provider_name))
```

```{r}
socialcare %>% 
  group_by(cic_type) %>% 
  summarise(observation = n(),
            location = n_distinct(location_id),
            provider = n_distinct(provider_name))
```

## logit models without COVID control 

```{r logit-sans-covid}
model_order_overall <- clm(rating ~ form + category + region + inherited,
                data = filter(socialcare, domain == "Overall"),
                link = "logit")
model_order_safe <- clm(rating ~ form + category + region + inherited,
                data = filter(socialcare, domain == "Safe"),
                link = "logit")
model_order_effective <- clm(rating ~ form + category + region + inherited,
                data = filter(socialcare, domain == "Effective"),
                link = "logit")
model_order_caring <- clm(rating ~ form + category + region + inherited,
                data = filter(socialcare, domain == "Caring"),
                link = "logit")
model_order_well_led <- clm(rating ~ form + category + region + inherited,
                data = filter(socialcare, domain == "Well-led"),
                link = "logit")
model_order_responsive <- clm(rating ~ form + category + region + inherited,
                data = filter(socialcare, domain == "Responsive"),
                link = "logit")
```

```{r}
ordinal_models <- modelsummary(list("overall" = model_order_overall, 
                                    "safe" = model_order_safe, 
                                    "effective" = model_order_effective, 
                                    "caring"= model_order_caring, 
                                    "well-led" = model_order_well_led, 
                                    "responsive" = model_order_responsive),
                               coef_omit = "region", exponentiate = F,
                               statistic = "({p.value}) {stars}")
ordinal_models
```


```{r}
ordinal_models_exp <- modelsummary(list("overall" = model_order_overall, 
                                        "safe" = model_order_safe, 
                                        "effective" = model_order_effective, 
                                        "caring"= model_order_caring, 
                                        "well-led" = model_order_well_led, 
                                        "responsive" = model_order_responsive),
                               coef_omit = "region", exponentiate = T,
                               statistic = "({p.value}) {stars}")
ordinal_models_exp
```

## logit models with COVID control 


```{r logit-covid-control}
model_order_overall_covid <- clm(rating ~ form + during_covid + 
                                   category + region + inherited,
                data = filter(socialcare, domain == "Overall"),
                link = "logit")
model_order_safe_covid <- clm(rating ~ form + during_covid +
                                category + region + inherited,
                data = filter(socialcare, domain == "Safe"),
                link = "logit")
model_order_effective_covid <- clm(rating ~ form + during_covid +
                                     category + region + inherited,
                data = filter(socialcare, domain == "Effective"),
                link = "logit")
model_order_caring_covid <- clm(rating ~ form + during_covid + 
                                  category + region + inherited,
                data = filter(socialcare, domain == "Caring"),
                link = "logit")
model_order_well_led_covid <- clm(rating ~ form + during_covid +
                                    category + region + inherited,
                data = filter(socialcare, domain == "Well-led"),
                link = "logit")
model_order_responsive_covid <- clm(rating ~ form + during_covid +
                                      category + region + inherited,
                data = filter(socialcare, domain == "Responsive"),
                link = "logit")
```


```{r}
ordinal_models_covid <-
  modelsummary(
    list(
      "overall" = model_order_overall_covid,
      "safe" = model_order_safe_covid,
      "effective" = model_order_effective_covid,
      "caring" = model_order_caring_covid,
      "well-led" = model_order_well_led_covid,
      "responsive" = model_order_responsive_covid
    ),
    coef_omit = "region",
    exponentiate = F,
    statistic = "({p.value}) {stars}"
  )
ordinal_models_covid
```

```{r}
ordinal_models_covid_exp <-
  modelsummary(
    list(
      "overall" = model_order_overall_covid,
      "safe" = model_order_safe_covid,
      "effective" = model_order_effective_covid,
      "caring" = model_order_caring_covid,
      "well-led" = model_order_well_led_covid,
      "responsive" = model_order_responsive_covid
    ),
    coef_omit = "region",
    exponentiate = T,
    statistic = "({p.value}) {stars}"
  )
ordinal_models_covid_exp
```

## OLS models with COVID control 

```{r}
# run the model with nest() method
# add after_covid as control
models_ols_covid <- socialcare %>% 
  mutate(domain = fct_relevel(domain, "Overall")) %>% 
  group_by(domain) %>% 
  nest()%>% 
  mutate(ols_models = map(data, 
                          ~lm(rating_num ~ form + during_covid + 
                              category + region + inherited, 
                              data = .x))) %>% 
  mutate(results = map(ols_models, ~ tidy(.x, conf.int = TRUE))) %>% 
  arrange(domain)


model_list <- models_ols_covid %>% 
  pull(ols_models, name = domain) 
```




```{r}
table_ols_covid <- modelsummary(model_list, statistic = "({p.value}) {stars}")
table_ols_covid
```

