---
title: "UK Social Care Data Cleanning"
author: "Meng Ye"
date: "8/12/2022"
output:
  pdf_document: default
  html_document: default
  word_document: default
---
## Note for update on 3/28/2024

1. change directory to relevant address
2. save both rds and csv
3. update cross-tab statistics to correspond to Table 1

## Introduction and Setup 

This file contains codes and steps that refine and merge the coded sampled CQC data, over-sampled (57 social care org) CQC data, coded individual/partnership providers data and also cross-check with Charity Commission (CC) data. 

I created this document as a stand-alone work file, which means I will start with *ONLY* raw data. And there will be some repeated code chunks from other document such as the one where we derived the CIC lists. 

We will recode the legal names to standardize their presentation, i.e. upper case, without ".", using "CIC" or "LTD", rather than the full expression. However, that new legal names will be saved in a new column, so that their original presentation will be saved should any merging with the skills for care data requires the exact way of presentation of organizations' legal name in their registrars' databases.

In the end, after running all the chunks, a comprehensive data set (our coded CQC sample to be analyzed) will be generated and saved in the end of the document. And all the codes for fitting models and running analyses will be written in a separate document. 

### Load packages 

```{r load package, message=FALSE, warning=FALSE}
library(tidyverse) # package for data cleaning and plotting
library(jsonlite) # to read json data
library(here) # manage directory
set.seed(5432)
```


### Import Raw Data 


```{r import data, message=FALSE, warning=FALSE}
# sampled data manually coded (with unfriendly column names)
socialcare_raw <- read_csv(here("data", "raw_data", "socialcare.csv"))

# Raw CQC full data (with unfriendly column names) to over sampling CICs
cqc_raw <- read_csv(here("data", "raw_data", "cqc_full.csv"))
 
# manually coded CIC list in the CQC data set, with CLS vs. CLG
cic_recoded <- read_csv(here("data", "raw_data", "cic_manual_recoded.csv"))

# import charities commission data
charities <- read_csv(here("data", "raw_data", "charities.csv"))

# manually coded (cross-checking against Companies House registration record) individuals and partnerships providers data
small <- read_csv(here("data", "raw_data", "small.csv"))
```


## Data cleanning - re-formating and preparation 


### 1a. Rename coded sample data frame columns to be more R-friendly 

```{r rename sample}
socialcare_renamed <- socialcare_raw %>% 
  rename(primary_cat = `Location Primary Inspection Category`,
         region = `Location Region`,
         domain = Domain,
         rating = latest_rating,
         inherited = `Inherited Rating (Y/N)`,
         provider_name = `Provider Name`,
         form_num = legal_form,
         publication_date = `Publication Date`) %>% 
  select(- `...1`, -`Location Type`) %>% 
  filter(location_type == "Social Care Org")
```

```{r}
head(socialcare_renamed)
```



### 1b. Standardize coded sample data frame's org name expression

```{r sample std}
socialcare_renamed <- socialcare_renamed %>% 
# remove dots
  mutate(std_name = str_replace_all(provider_name, "\\.", "")) %>% 
# capitalize all
  mutate(std_name = str_to_upper(std_name)) %>%
# standardize to LTD
  mutate(std_name = str_replace_all(std_name, "LIMITED", "LTD")) %>% 
# standardize to CIC
  mutate(std_name = str_replace_all(std_name, "COMMUNITY INTEREST COMPANY", "CIC"))
```


### 2a. Rename full cqc data frame columns to be more R-friendly 

```{r rename cqc full data set}
cqc_renamed <- cqc_raw %>% 
  rename(primary_cat = `Location Primary Inspection Category`,
         region = `Location Region`,
         domain = Domain,
         rating = `Latest Rating`,
         inherited = `Inherited Rating (Y/N)`,
         provider_name = `Provider Name`,
         location_id = `Location ID`,
         location_type = `Location Type...26`,
         publication_date = `Publication Date`) %>% 
  select(-`Location Type...5`, -`Legal Form`) %>% 
  filter(location_type == "Social Care Org")


```


### 2b. Standardize coded sample data frame's org name expression

```{r cqc std}
cqc_renamed <- cqc_renamed %>% 
# remove dots
  mutate(std_name = str_replace_all(provider_name, "\\.", "")) %>% 
# capitalize all
  mutate(std_name = str_to_upper(std_name)) %>%
# standardize to LTD
  mutate(std_name = str_replace_all(std_name, "LIMITED", "LTD")) %>% 
# standardize to CIC
  mutate(std_name = str_replace_all(std_name, "COMMUNITY INTEREST COMPANY", "CIC")) %>% 
# standardize to CIO
  mutate(std_name = str_replace_all(std_name, "CHARITABLE INCORPORATED ORGANISATION", "CIO")) 
```


### 3. Standardize CIC list org name expression 

```{r cic full list std}
cic_renamed <- cic_recoded %>% 
  select(-`...1`, - legal_name, cic_number)  %>% 
# remove dots
  mutate(std_name = str_replace_all(provider_name, "\\.", "")) %>% 
# capitalize all
  mutate(std_name = str_to_upper(std_name)) %>%
# standardize to LTD
  mutate(std_name = str_replace_all(std_name, "LIMITED", "LTD")) %>% 
# standardize to CIC
  mutate(std_name = str_replace_all(std_name, "COMMUNITY INTEREST COMPANY", "CIC")) %>% 
# standardize to CIO
  mutate(std_name = str_replace_all(std_name, "CHARITABLE INCORPORATED ORGANISATION", "CIO")) %>% 
# delete unnecessary columns to make sure combining row align better later
  select(- cic_number, - provider_name)
```


### 4a. Derive a smaller sub set (select necessary columns) of Charity Commision data

```{r CC subset}
charity_selected <- charities %>% 
# filter out only active charities
  filter(charity_registration_status == "Registered") %>% 
# select only potentially useful columns 
  select(charity_number = registered_charity_number,
         charity_name, charity_type,
         charity_registration_date = date_of_registration,
         charity_company_number = charity_company_registration_number)
```


### 4b. Standardize Charity Commision org name expression 

```{r CC std}
charity_renamed <- charity_selected %>% 
# remove dots
  mutate(std_name = str_replace_all(charity_name, "\\.", "")) %>% 
# capitalize all
  mutate(std_name = str_to_upper(std_name)) %>%
# standardize to LTD
  mutate(std_name = str_replace_all(std_name, "LIMITED", "LTD")) %>% 
# standardize to CIC
  mutate(std_name = str_replace_all(std_name, "COMMUNITY INTEREST COMPANY", "CIC")) %>% 
# standardize to CIO
  mutate(std_name = str_replace_all(std_name, "CHARITABLE INCORPORATED ORGANISATION", "CIO")) 
```


### 5.Standardize individual/partnership provider name expression 

Based on Dr. Kerlin's code, this is what I did:


1.	Leave it as the same type individual/partnerships for blank rows
2.  Leave it as the same type individual/partnerships for rows with No-No for both Charity Commission and Company House
3.	Recode those have limited by shares Company House records as “for-profit”. 
4.	Recode “LLP” to for-profit 
5.  Recode company limited by guarantee to "NA", because we are not sure if it is nonprofit or not
(See https://en.wikipedia.org/wiki/Private_company_limited_by_guarantee, especially under Commercial enterprises)
4.	Recode to “NA” for those marked with several possibilities 


```{r rename indivial and partnership}
small_renamed <- small %>% 
  select(provider_name, small_form_num) %>% 
  
# remove dots
  mutate(std_name = str_replace_all(provider_name, "\\.", "")) %>% 
# capitalize all
  mutate(std_name = str_to_upper(std_name)) %>%
# standardize to LTD
  mutate(std_name = str_replace_all(std_name, "LIMITED", "LTD")) %>% 
# standardize to CIC
  mutate(std_name = str_replace_all(std_name, "COMMUNITY INTEREST COMPANY", "CIC")) %>% 
# standardize to CIO
  mutate(std_name = str_replace_all(std_name, "CHARITABLE INCORPORATED ORGANISATION", "CIO")) %>% 
  select(-provider_name)
```



## Data cleanning - merging and recoding 

### 1. Use full CIC list and complete CQC data to derive full CIC data frame

```{r cic over sample data}
cic_over_sample <- cic_renamed %>% 
  left_join(cqc_renamed, by = "std_name") %>% 
# mark that the form is cic
  mutate(form_num = 4)
```

Now we check the merged data, there are 540 rows (for 57 organizations because of multiple locations and multiple rating domains) and 33 variables. 
```{r cic over sample data check}
head(cic_over_sample, 10)
```


### 2. Merge orginical sample with the full CIC sample 

First we carve out the CICs from the old samples 
```{r merge old sample and cic}
sample_no_cic <- socialcare_renamed %>% 
## leave out old cic orgs from the coded sample
  filter(form_num != 4)
```

Then we combine the sample without (old) CICs and the full CICs data

```{r add full cic}
sample_v2 <- bind_rows(sample_no_cic, cic_over_sample)
head(sample_v2)
```


### 3. Merge the updated sample v2 with indivial/partnership data

```{r merge in small}
# select fewer columns to merge
sample_v3 <- sample_v2 %>% 
# add data with refined coding
  left_join(small_renamed, by = "std_name") %>% 
# supersede old coding with new coding
  mutate(form_num = ifelse(form_num == 5, small_form_num, form_num)) %>% 
  select(- small_form_num)

tail(sample_v3)
```

### 4. Merge the updated sample v3 with Charity data

```{r}
# select charity std names and code the form type as "2".
charity_list <- charity_renamed %>% 
  select(std_name) %>% 
  mutate(charity_form_num = 2) %>% 
# Drop repetition
  group_by(std_name) %>% 
  slice(1) %>% 
  ungroup()

sample_v4 <- sample_v3 %>% 
# add data with refined coding
  left_join(charity_list, by = "std_name") %>% 
# supersede old coding with new coding
  mutate(form_num = ifelse(is.na(charity_form_num), form_num, charity_form_num)) %>% 
  select(- charity_form_num)

sample_new <- sample_v4%>% 
  mutate(index = 1:nrow(sample_v4))
```


## Check how much change was made


```{r distribution old name}
sample_new %>% 
  count(form_num)
```


Note:
                case_when(form_num == 1 ~ "FPO",
                          form_num == 2 ~ "NPO",
                          form_num == 3 ~ "GOV",
                          form_num == 4 ~ "CIC",
                          form_num == 5 ~ "IND",


## Save data

```{r eval = FALSE}
write_rds(sample_new, file = here("data", "derived_data", "sample_new.rds"))
write_csv(sample_new, file = here("data", "derived_data", "sample_new.csv"))
```





```{r}
count(sample_new, factor(form_num), domain)
```

```{r count-by-level}
sample_new %>% 
  mutate(overall = ifelse(domain == "Overall", 1, 0)) %>% 
  summarize(count_provider = n_distinct(provider_name),
            count_location = n_distinct(location_id),
            count_overall_rating = sum(overall),
            count_rating = n())
```

```{r count-by-form-level}
# counts of multi-level observations by the form types
# statistics corresponding to Table 1
sample_new %>% 
  mutate(overall = ifelse(domain == "Overall", 1, 0)) %>% 
  group_by(factor(form_num)) %>% 
  summarize(count_provider = n_distinct(provider_name),
            count_location = n_distinct(location_id),
            count_overall_rating = sum(overall),
            count_rating = n())
```


Note:
                case_when(form_num == 1 ~ "FPO",
                          form_num == 2 ~ "NPO",
                          form_num == 3 ~ "GOV",
                          form_num == 4 ~ "CIC",
                          form_num == 5 ~ "IND",
                          
                          
                          
                          
